A benchmark for the implementation of standard vehicle routing algorithms

### 2021-02-18 paper/main.tex
paper progress

\title{Benchmark!!1}
\author{
Fabien Tricoire$^{1}$ \\[1ex]
 \small $^1$Institute for Transport and Logistics Management,
 Vienna University of Economics and Business\\
 \small Welthandelsplatz 1, 1020 Vienna, Austria\\
  \small \texttt{fabien.tricoire@wu.ac.at}\\[2ex]
}
\date{}
\maketitle

@@

\section{The benchmarks}
We consider five different benchmark algorithms that perform tasks
commonly encountered in transport optimization. Together they capture
key aspects of computationally intensive tasks in heuristics and
exact methods for transport optimization, although they do not
necessarily use the most efficient algorithm to solve the problem they
are tackling. Additionally, we believe
that they also capture such aspects for other fields of application
of operations research, such as e.g. scheduling. For the sake of
simplicity, we consider the symmetric travelling salesperson problem
(TSP) as a base problem, i.e. input data are in the form of a
symmetric distance matrix while a solution is a permutation. However
some benchmark algorithms solve different problems than the TSP based on these
TSP data, as described below.

@@

For each benchmark the reference CPU time is provided by the c++14
implementation. We also develop a variant of c++14 using a nested
distance matrix, and another using static arrays.

@@

\subsection{Julia}
Julia is a relatively recent programming language. Mention dates for
first release (0.4?) as well as 1.0.

Flat vs square matrix, mention 2-dimensional arrays.

\subsection{Rust}

\subsection{JavaScript}
JavaScript is not expected to perform as well as, say, C++. However
there are several reasons why it can be appealing to implement routing
algorithms in JavaScript:
\begin{itemize}
\item Any program written in JavaScript can run on virtually any
  computer with a modern web browser, including smartphones.
\item Integrating JavaScript with a web interface is especially easy.
\item JavaScript engines have received considerable attention from
  major companies and are the subject of fierce competition on
  performance. As a result they have seen vast performance
  improvements over the years and the trend is likely to continue.
\end{itemize}

@@

Flat vs nested.

@@

Mention here compilers and compiler versions used.

@@

>>>>

\title{A benchmark for the implementation of standard vehicle routing
  algorithms}
\author{
Fabien Tricoire$^{1}$ \\[1ex]
 \small $^1$Institute for Transport and Logistics Management,
 Vienna University of Economics and Business\\
 \small Welthandelsplatz 1, 1020 Vienna, Austria\\
  \small \texttt{fabien.tricoire@wu.ac.at}\\[2ex]
}
\date{}
\maketitle

@@

\section{The benchmarks}
\label{sec:benchmark}
We consider five different benchmark algorithms that perform tasks
commonly encountered in transport optimization. Together they capture
key aspects of computationally intensive tasks in heuristics and
exact methods for transport optimization, although they do not
necessarily use the most efficient algorithm to solve the problem they
are tackling. Additionally, we believe
that they also capture such aspects for other fields of application
of operations research, such as e.g. scheduling. For the sake of
simplicity, we consider the symmetric travelling salesperson problem
(TSP) as a base problem, i.e. input data are in the form of a
symmetric distance matrix while a solution is a permutation. However
some benchmark algorithms solve different problems than the TSP based on these
TSP data, as described below.

@@

For each benchmark the reference CPU time is provided by the c++98
implementation. We also develop a variant of c++14 using a nested
distance matrix, and another using static arrays.

@@

\subsection{Julia}
Julia is a relatively recent programming language. It is a
general-purpose language but it is also particularly aimed at
computationally intensive tasks such as numerical
analysis~\cite{Julia-2017}. Version 1.0 was 
released in 2018 and the language is well-documented and has a strong
community.

Compilation happens in a JIT fashion. Like Numpy but
unlike all other programming languages considered, Julia supports
multi-dimensional arrays, therefore a native two-dimensional array can be
used to represent a distance matrix. We also develop an
implementation using flat matrix representation in Julia, called
\emph{julia-flat-matrix}; we will determine which version is better
through experiment.

One valuable feature of Julia is the fact that projects have an
environment, including package dependencies with specific
versions. The language makes it easy to \emph{instantiate} the
project environment, i.e. recreate the same environment in which the
project was developed, with the same packages in the same
versions. This provides safety from new library versions which break
backward compatibility.

\subsection{Rust}
Rust is also relatively recent, with version 1.0 released in 2015. It
is a general-purpose language with an emphasis on memory safety and
on performance. Notable software using Rust include Firefox, Dropbox
and Cloudflare~\cite{rust-lang}.

Programs in Rust are pre-compiled, like in C++ for
instance. Like Julia, the language is well documented and has a strong
community. Also like Julia, projects have an environment, and it is
easy to recreate the environment for a given project, including
dependencies and their specific version, which again provides safety
from backward-compatibility-breaking library updates. The \emph{rust}
implementation uses a flat matrix representation.

\subsection{JavaScript}
JavaScript is likely one of the most popular programming languages in
the World, as it is used on countless websites as well as mobile
applications based on web technology. JavaScript is not expected to
perform as well as, say, C++. However there are several reasons why it
can be appealing to implement routing algorithms in JavaScript:
\begin{itemize}
\item Any program written in JavaScript can run on virtually any
  computer with a modern web browser, including smartphones.
\item Integrating JavaScript with a web interface is especially easy.
\item JavaScript engines have received considerable attention from
  major companies and been the subject of fierce competition in
  relation with web browsers engines. As a result they have seen vast
  performance improvements over the years and the trend is likely to continue.
\end{itemize}

@@

The \emph{javascript} implementation uses a flat matrix representation
but there is also an implementation using nested matrix representation,
called \emph{javascript-flat-matrix}.

@@

A summary of all versions used for compilers and interpreters is
provided in Appendix~\ref{app:lang}.

\subsection{Benchmark data}
Each instance file contains one $n \times n$ distance matrix as well
as $p$ randomly generated permutations, which are used as starting
\emph{seeds} for the algorithms (e.g. as starting solution for 2-opt).
For $n$ taking each value of $\{20, 40, 60\}$ and fixing $p = 1000$ we
generate 40 instance files, for a total of 120 instance files. Each
instance file contains 1000 permutations, therefore is run 1000
different times for each benchmark. As mentioned in
Section~\label{sec:benchmark} distances are Euclidean distances multiplied
by 100 and truncated to their integer part. In the following each data
point represents the total CPU time for running one implementation
for each of the 1000 permutations of one given instance file.

@@

\section{Summary of language and compiler versions}
\label{app:lang}
\begin{table}
  \begin{tabular}{l|l}
    \toprule
    Language/tool & language or compiler version & options \\
    \midrule
    C++14 & g++ & -Wall -ansi -pedantic -O3 -std=c++14 \\
    C++98 & g++ & -Wall -ansi -pedantic -O3 -std=c++98 \\
    Python & & \\
    Pypy & & \\
    Numpy & & \\
    Numba & & \\
    Java &  & \\
    Julia & 1.5.3 & --check-bounds=no --inline=yes -O3 -t 1 \\
    Rust & & \\
    JavaScript & Node & \\
    \bottomrule   
  \end{tabular}
  \caption{Overview of all implementations}
\end{table}
